---
title: "How to calculate confidence intervals"
output:
  html_notebook: default
  html_document:
    df_print: paged
  word_document: default
---

Consider our snow vole case study... We will calculate the confidence intervals for the mean of the weight of the individuals, for the foot length, and for different groups. 

```{r packages}
library(Rmisc)
library(Hmisc)
library(ggplot2)
library(boot)
```

```{r load-data}
dati <- read.csv("data/captures.csv", sep=";")
```


# Calculating and plotting CIs with the traditional method

The function `CI` in the `Rmisc` package produces confidence intervals for the mean using the traditional method:
si deve specificare la colonna di dati che vogliamo considerare
```{r}
CI(dati$weight_g, ci=0.95)
```

So what?? How can we fix that??
puliamo il dataset...NA va usato con cautela perchè in alcuni casi
se non correttamente impostato ci cancellerebbe tutti i dati di tutte le 
colonne

```{r}
CI(na.omit(dati$weight_g), ci=0.95)
```

What about the CIs for different group of animals?
la tilde vuol dire "tenendo conto del sesso degli animali"
```{r}
dati2 <- na.omit(dati)
group.CI(weight_g ~ sex,
         data= dati2,
         ci=0.95)
intervalli <- group.CI(weight_g ~ sex,
         data= dati2,
         ci=0.95)
intervalli
```
domanda legittima; è ragionevole che le femmine siano mediamente più pesanti dei maschi?
se nell'esercizio precedente abbiamo riscontrato la conclusione opposta
femmine gravide?
```{r}
ggplot(intervalli, aes(x=sex, y=weight_g.mean, colour=sex)) +
    geom_errorbar(aes(ymin=weight_g.lower, ymax=weight_g.upper), width=.1) +
    geom_line() +
    geom_point()
```

Do it on your own:

* plot just one of the confidence intervals (for males or for females, of for the overall mean)
* plot the confidence intervals for the mean weights, grouped for age
* repeat for the footlength


```{r}
as.data.frame(CI(na.omit(dati$weight_g), ci=0.95)) -> CImean
CImean
```

Footlength CI:

```{r}

variabile <- dati2$footlength_mm # o qualsiasi altra variabile numerica

# script 'generalizzato'
intervalli <- group.CI(variabile ~ sex,
         data= dati2,
         ci=0.95)
intervalli
ggplot(intervalli, aes(x=sex, y=variabile.mean, colour=sex)) +
    geom_errorbar(aes(ymin=variabile.lower, ymax=variabile.upper), width=.1) +
    geom_line() +
    geom_point()
```
le femmine sembrano avere un piede più piccolo dei maschi, e minore ampiezza
degli intervalli di confidenza. Visti questi risultati possiamo fare un test
di ipotesi.

# Calculating and plotting CIs with the bootstrapping method

We first need to apply the bootstrap to our data. For this purpose, we use the function `boot` in the `boot` package:

```{r}
boot.data <- boot(dati2$weight_g,
                  function(x,i) mean(x[i]), 
                  R=10000)
```
R=10000 è il numero di repliche di campionamento.
Now we can use the bootstrapped data to calculate the CIs, using the `boot.ci` function:

```{r}
boot.ci(boot.data,
        conf = 0.95)
```

The function generates different types of non-parametric confidence intervals. For routine use, it is recommended to use the BCa or percentile methods. The formulae on which the calculations are based can be found in Chapter 5 of Davison and Hinkley (1997). Davison, A.C. and Hinkley, D.V. (1997) Bootstrap Methods and Their Application, Chapter 5. Cambridge University Press.

Do it on your own:

* Calculate the bootstrapped intervals for the sex groups
* Plot the bootstrapped intervals

adesso possiamo calcolare gli intervalli di bootstrap separatamente
per i maschi e per le femmine
```{r}
library(tidyverse)
dati2 %>%
  filter(sex == "F") -> datiF
dati2 %>%
  filter(sex == "M") -> datiM

dataset <- datiM

boot.data <- boot(dataset$weight_g,
                  function(x,i) mean(x[i]), 
                  R=10000)
boot.ci(boot.data,
        conf = 0.95)
```

## Optional readings:   

* 'Calculating the Confidence interval for a mean using a formula' from Statistics Learning Center. (Dr. Nic). 2013. www.youtube.com/watch?v=s4SRdaTycaw
* Davison, A.C. and Hinkley, D.V. (1997) Bootstrap Methods and Their Application, Chapter 5. Cambridge University Press.

