---
title: 'Two sample t-test example #2'
output:
  html_document:
    df_print: paged
---
valutazione di un test svolto da più studenti
# Problem - dilemma of two graders

In order to promote fairness in grading, each application was graded twice by different graders. Based on the grades, can we see if there is a difference between the two graders? The data is

```{r}
x <- c(3, 0, 5, 2, 5, 5, 5, 4, 4, 5)
y <- c(2, 1, 4, 1, 4, 3, 3, 2, 3, 5)
```
i due vettori non sono più indipendenti, è lo stesso compito che viene valutato
da due persone diverse. I due valori dei vettori sono riferiti alla stessa
unità sperimentale che stiamo considerando.
Clearly there are differences. Are they described by random fluctuations, or is there a bias of one grader over another? 
Verify the t-test assumptions, and then proceed with the t-test if appropriate. Which is the corresponding non-parametric test? Perfom it and compare the results to the one obtained with the t-test.

# Solution

```{r}
# normality test

# with a qqplot
par(mfrow=c(1,2))
qqnorm(x); qqline(x, col="red")
qqnorm(y); qqline(y, col="red")

# with the Shapiro-Wilk test
shapiro.test(x)
shapiro.test(y)
```


```{r}
# homogeneity of variances
var.test(x,y)
```
la novità è paired=true
```{r}
# t-test
#t-test
t.test(x,y,
       paired = TRUE,
       var.equal = TRUE,
       alternative = "two.sided")
```

```{r}
# non parametric test
wilcox.test(x,y,
            paired = TRUE,
            alternative = "two.sided")
```
