---
title: "Factorial experiment example - weight gain"
output: 
  html_document: 
    theme: paper
    toc: yes
    toc_float: yes
  html_notebook: 
    toc: yes
    toc_float: yes
---

# Problem
(source: Crawley, The R book)

Our example comes from a farm-scale trial of animal diets.
There are two factors: diet and supplement.

Diet is a factor with three levels: barley, oats and wheat.

Supplement is a factor with four levels: agrimore, control, supergain and supersupp.

The response variable is weight gain after 6 weeks.

```{r}
rm(list = ls())
weights <- read.csv("data/growth2.txt", sep="\t")
head(weights)
```

```{r}
table(weights$diet, weights$supplement)
```
la funzione table ci conta quante volte barley è associata ad agrimore e cos' via....il campione è bilanciato.

```{r}
with(weights,
     coplot(gain ~ diet|supplement,rows=2)) #aggiusto, per ora tralalsciamo
```
fittiamo l'analisi della varianza inserendo il termine di interazione.
We will perform an analysis of variance including a factor for the interaction between diet and supplement:

```{r}
model<-aov(gain~diet*supplement, data = weights)
summary(model)
```
non c'è un effetto di interazione tra la dieta e l'integratore (supplemento), anche se c'è almeno una dieta che funziona meglio delle altre.
The ANOVA table shows that there is no hint of any interaction between the two
explanatory variables (p=0.917): the effects of diet and supplement are additive.    

Let's check for this graphically...

```{r}
weights$diet <-as.factor(weights$diet)
weights$diet <- relevel(weights$diet, ref="oats")
with(weights, interaction.plot(diet,supplement, response = gain))
```
abbiamo un picco in corrispondenza della dieta barley, indipendente dall'integratore, inoltre le rette sono quasi parallele tra loro. Se le interazioni fossero significative le rette si intersecherebbero tra di loro.

Since interactions are not significant, we can consider simplifying our model:
togliamo il termine di interazione poichè abbiamo visto non è significativa.
```{r}
model2 <- aov(gain~diet + supplement, data = weights)
summary(model2)
```

Which one is best?

The second model is much simpler than the first one:
"restituiscimi l'output dell'aNOVa come se avessi fatto un linear model"
```{r}
summary.lm(model)
```


```{r}
summary.lm(model2)
```
nel secondo modello , più semplice ho dimezzato il numero di parametri. Per decidere quale dei modelli è significativamente migliore è meglio confrontare le varianze
We can compare them directly:
```{r}
anova(model,model2)
```
risultato; non c'è una differenza significativa di varianza tra i due modelli (pf 0.9166). Quindi "possiamo scegliere" il modello 2.
The second model is simpler and it is not significantly worst than the first model, so we go for model2 and we try to interpret it.

Go back and look again at the summary of model2:   

* the barley and wheat diet are statistically different, in terms of weight gain, from the oats diet, but what about the difference between wheat and barley?
* it's the same for the control and supergain supplements with respect to agrimore, but what about the difference between them? and with supersupp??

Let's check with a TukeyHSD test:

```{r}
TukeyHSD(model2)
```
guardiamo le differenze tra le diete. ci dà la conferma che barley è diverso da oaths e così per tutte le altre coppie.

```{r}
plot(TukeyHSD(model2))
```

Other plots using the jtools package:
```{r}
library(sjPlot)
# plot_model(model2, type = "diag")
plot_model(model2, type="eff", terms=c("diet"))
```

check here for an example with a significant interaction: https://cran.r-project.org/web/packages/jtools/vignettes/categorical.html


# Interactions with at least one continuous predictor
(source: https://cran.r-project.org/web/packages/jtools/vignettes/interactions.html)


This famous (Fisher's or Anderson's) iris data set gives the measurements in centimeters of the variables sepal length and width and petal length and width, respectively, for 50 flowers from each of 3 species of iris. The species are Iris setosa, versicolor, and virginica.

```{r}
data(iris)
head(iris)
```

Please note that here we use `lm` rather than `aov`:

```{r}
fitiris <- lm(Petal.Length ~ Petal.Width * Species, data = iris)
library(jtools)
plot_model(fitiris, type="int")
```

nella specie "setosa" la lunghezza del petalo aumenta gradualmente con la larghezza del petalo, con versicolor l'aumento è più netto.