---
title: "Multicollinearity Essentials and VIF in R"
output: 
  html_document: 
    theme: paper
    toc: yes
    toc_float: yes
  html_notebook: 
    toc: yes
    toc_float: yes
---


```{r}
library(MASS)
library(car)
library(dplyr)
library(ggcorrplot)
```


# Collinearity

[Source](http://www.sthda.com/english/articles/39-regression-model-diagnostics/160-multicollinearity-essentials-and-vif-in-r/) 

In multiple regression, two or more **predictor variables** might be correlated with each other. This situation is referred as **collinearity**.

There is an extreme situation, called **multicollinearity**, where collinearity exists between three or more variables even if no pair of variables has a particularly high correlation. This means that there is **redundancy** between predictor variables.

In the presence of multicollinearity, **the solution of the regression model becomes unstable**.


# Example

## Correlation analysis

Collinearity issues can be assessed in different ways. The simplest one involves computing the correlation coefficients between variables.

We’ll use the Boston data set in `MASS` package for predicting the median house value (`mdev`) in Boston Suburbs, based on multiple predictor variables.
Il dataset Boston contiene tutta una serie di variabili x, noi stiamo cercando di spiegare il prezzo delle case in una certa zona, e poi abbiamo tutte le caratteristiche di quella certa zona.
```{r}
# Load the data
data("Boston", package = "MASS")
# Split the data into training and test set
# set.seed(123)
# training.samples <- Boston$medv %>%
#   createDataPartition(p = 0.8, list = FALSE)
# train.data  <- Boston[training.samples, ]
# test.data <- Boston[-training.samples, ]
```

We look at the data:

```{r}
head(Boston)
```

We select the predictor variables:

```{r}
pred <- dplyr::select(Boston, crim:lstat)
names(pred)
pred
```

Now, we can compute a correlation matrix:
utilizziamo la funzione cor per creare una matrice di correlazione.
```{r}
cor.matrix <- cor(pred)
cor.matrix
```
ad esempio, tra tax e indus c'è un coefficiente di 0.7 abbastanza alto che già ci dice qualcosa.
And we can visualize it:
se il cor>0.7, eliminiamo una variabile.PEr scegliere quale, consideriamo il significato della variabile, la facilità di reperibilità dei dati per quella variabile, e se non riuscissimo a scartare neanche così facciamo un modello,e lasciamo alla matematica decidere il modello migliore che "fitta" meglio, in base all'AIC appunto.
```{r}
ggcorrplot(cor.matrix)
ggcorrplot(cor(pred))
```

Just a different visualization:

```{r}
ggcorrplot(cor.matrix, method = "circle")
```

We can add labels, but it is still difficult detecting relevant correlations: 

```{r}
ggcorrplot(cor.matrix, method = "circle", lab=TRUE)
```
possiamo chiedere a R di andare a identificare nella nostra matrice dove ci sono i problemi di correlazione. Prendiamo in valore assoluto la matrice precedente, perchè appunto non importa il segno, e andiamo a dirgli di metterci un TRuE o FALSE.
We set a threshold to identify relevant correlations:

```{r}
p.mat <- (abs(cor.matrix) > 0.7)
p.mat
```
ora, R ci mette una crocetta su quelle coppie di variabili che sono correlate tra loro . Ci sono 7 coppie di variabili correlate.

```{r}
# Argument p.mat
# Barring the significant coefficients
ggcorrplot(cor.matrix,
  hc.order = TRUE, method = "circle", #metti in ordine i valori
  type = "lower", p.mat = p.mat #stampami solo la parte inferiore della matrice
) #perchè è una matrice speculare.
```
Indus ha ben tre problemi di correlazione, quindi posso agire per eliminarlo, perchè è molto correlato. Posso procedere tenendo quelle che non sono correlate insieme a quelle da tenere con attenzione.Di solito per ogni parametro stimato dobbiamo avere almeno 10-15 righe nel dataset.
questo era il primo metodo "a monte".....poi possiamo fqare un modello lineare che consideri tutte le variabili e mi dia il suo summary.
When faced to multicollinearity, the concerned variables should be removed, since the presence of multicollinearity implies that the information that this variable provides about the response is redundant in the presence of the other variables (James et al. 2014,P. Bruce and Bruce (2017)).

## Variance Inflation Factor

For a given predictor, multicollinearity can also be assessed by computing a score called the variance inflation factor (or **VIF**), which **measures how much the variance of a regression coefficient is inflated due to multicollinearity in the model**.

The smallest possible value of VIF is **one (absence of multicollinearity)**. As a rule of thumb, **a VIF value that exceeds 5 or 10 indicates a problematic amount of collinearity (James et al. 2014)**.

The following regression model include all predictor variables:

```{r}
# Build the model
model1 <- lm(medv ~ ., data = Boston)
summary(model1)
```
non possiamo fidarci di questo output perchè l'errore standard di 3.8 "non ci possiamo credere" (è un esempio) perchè la multicollinearità fa cambiare le stime , perchè prima dall'analisi precedente avevamo ottenuto alcune correlazioni>0.7.

The R function `vif()`, (variance inflation factor)in the `car` package, can be used to detect multicollinearity in a regression model:

```{r}
vif(model1)
```
Generalmente se il VIF è maggiore di 5/10 ci indica che quelle variabili stanno soffrendo la collinearità e andrebbero eliminate dal dataset. 
In this case, we can see that the `rad` and `tax` predictors have high VIF values - compare with the results of the correlation analysis above!!

# Exercise

Consider a dataset from air pollution studies. How is sulphur dioxide concentration (`Pollution`) related to other variables in the dataset? Are there any problems of collinearity?

```{r}
pollution <- read.table("data/sulphur.dioxide.txt", header=TRUE)
View(pollution)
```

In multiple regression, it is always a good idea to use pairs to look at all the correlations:

```{r}
pairs(pollution,panel=panel.smooth)
```
generiamo degli "scatterplot a coppie per la nostra matrice di correlazione"
Check for relevant correlations and visualize them. Compute the variance inflation factor and update the model accordingly.

```{r}
cor.matrix <- cor(pollution)
# cor.matrix
p.mat <- (abs(cor.matrix) > 0.7)
# p.mat
ggcorrplot(cor.matrix,
  hc.order = TRUE, method = "circle",
  type = "lower", p.mat = p.mat
)
```
su industria e popolazione c'è un caso effettivamente problematico.
```{r}
# Build the model
model1 <- lm(Pollution ~ ., data = pollution)
summary(model1)
```
```{r}
vif(model1)
model1 <- lm(Pollution ~ Temp + Industry + Wind + Rain + Wet.days, data = pollution)
summary(model1)
```
togliendo la variabile population si è abbassato drasticamente l'errore standard su Industry.
solo dopo aver calcolato il VIF ci concentriamo sull'output. Solo alla fine, ed escludendo ogni ipotesi di collinearità si valutano gli AIC tra i modelli.
